---
title: "Class 5 - ggplot"
author: "Benjamin Lee"
format: pdf
---

# Our first ggplot

To use the ggplot2 package I first need to have it installed on my computer

To install any package we use the 'install.packages()' command.

Now can I use it? NO ! first we need to call 'library(ggplot2)'
```{r}
library(ggplot2)
ggplot()
```
```{r}
mpg
```

Our first plot of displ vs hwy 
All ggplot() graphs are made in the same way. 

- data + aes + geoms
```{r}
ggplot(mpg) +
  aes(x=displ, y=hwy) + 
  geom_point() 
```

I can add more layers: 
```{r}
ggplot(mpg) +
  aes(x=displ, y=hwy) + 
  geom_point() +
  geom_smooth(method = lm, se = FALSE)
```

# Plot of gene expression data

First read the data from online. 

```{r}
url <- "https://bioboot.github.io/bimm143_S20/class-material/up_down_expression.txt"
genes <- read.delim(url)
head(genes)
```

>Q. How many genes are in this dataset?

```{r}
colnames(genes)
```

What are the colnames?
```{r}
colnames(genes)
```

A first version of this data Condition1 vs Condition2
```{r}
ggplot(genes) + 
  aes(x=Condition1 , y=Condition2) +
  geom_point()
```

Let's add some color. To do this we can add another aes() mapping of color the 'State' column in our data.
```{r}
ggplot(genes) + 
  aes(x=Condition1 , y=Condition2, col = State) +
  geom_point()
```

> Q. How many genes are upregulated and downregulated?

```{r}
head(genes)
```

To get at just the State column
```{r}
table(genes$State)
```


```{r}
p <- ggplot(genes) + 
    aes(x=Condition1, y=Condition2, col=State) +
    geom_point() 
```
```{r}
p + scale_colour_manual( values=c("blue","gray","red") )
```

```{r}
ggplot(genes) + 
    aes(x=Condition1, y=Condition2, col=State) +
    geom_point() + 
  scale_colour_manual( values=c("blue","gray","red") )
```
Then just add to our object 'p'
```{r}
p + scale_colour_manual( values=c("blue","gray","red") ) + 
  labs(title="Gene Expresion Changes Upon Drug Treatment",
         x="Control (no drug) ",
         y="Drug Treatment")
```
