---
title: "class 10 - Halloween mini project"
author: "Benjamin Lee"
format: pdf
---

```{r}
candy_file <- "https://raw.githubusercontent.com/fivethirtyeight/data/master/candy-power-ranking/candy-data.csv"

candy = read.csv(candy_file, row.names=1)
head(candy)

```
> Q1. how many different candy types are in this dataset?

```{r}
nrow(candy)
```

> Q2. How many fruity candy types are in the dataset?

```{r}
sum(candy$fruity)
```

## Winpercent 
The most interesting variable in the dataset is 'winpercent'. for a given candy this value is the percentage of people who prefer this candy over another randomly chosen candy 

> Q3. What is your favorite candy in the dataset and what is it’s winpercent value?

```{r}
rownames(candy)
```
```{r}
candy["Sour Patch Kids", ]$winpercent

```

> Q4. What is the winpercent value for “Kit Kat”?

```{r}
candy["Kit Kat",] $winpercent
```

> Q5. What is the winpercent value for “Tootsie Roll Snack Bars”?

```{r}
candy["Tootsie Roll Snack Bars",]$winpercent
```

```{r}
library("skimr")
skim(candy)
```

> Q8. Plot a histogram of winpercent values

```{r}
hist(candy$winpercent)
```

```{r}
library(ggplot2)
ggplot(candy) +
  aes(winpercent) + 
  geom_histogram(bins = 10, col ="red", fill="orange")
```

> Q11. On average is chocolate candy higher or lower ranked than fruity candy?

```{r}
chocolate.inds <- as.logical(candy$chocolate)
choc.win <- candy[chocolate.inds,]$winpercent

fruity.inds <- as.logical(candy$fruity)
fruity.win <- candy[fruity.inds,]$winpercent

mean(choc.win)
mean(fruity.win)
```

> Q12. Is this difference statistically significant?

```{r}
t.test(choc.win, fruity.win)

```

# 3. Candy ranking 

First setup some colors for different candy types

```{r}
my_cols = rep("black", nrow(candy))
my_cols[as.logical(candy$chocolate)] = "chocolate"
my_cols[as.logical(candy$bar)] = "brown"
my_cols[as.logical(candy$fruity)] = "pink"
my_cols
```

```{r}
ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy), winpercent)) +
  geom_col(fill = my_cols)
```
```{r}
ggsave("tmp.png")
```

Now, for the first time, using this plot we can answer questions like:
> Q17. What is the worst ranked chocolate candy?
> Q18. What is the best ranked fruity candy?


##4. Taking a look at pricepercent

What is the best (most liked in terms of 'winpercent') for the money (in terms of 'pricepercent')?

To answer this I will make a plot of winpercent vs pricepercent
```{r}
library(ggrepel)

# How about a plot of price vs win
ggplot(candy) +
  aes(winpercent, pricepercent, label=rownames(candy)) +
  geom_point(col=my_cols) + 
  geom_text_repel(col=my_cols, size=3.3, max.overlaps = 5)
```

## 5. Exploring the correlation structure
```{r}
library(corrplot)
cij <- cor(candy)
corrplot(cij)
```

